This is an example of article title

### 2025-11-05 paper/paper.tex
Fixing a few minor LaTeX quirks and typos in the JuliaCon paper.

% JuliaCon proceedings template
\documentclass{juliacon}
\setcounter{page}{1}
\usepackage{amsmath,booktabs,mathtools,amssymb}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{balance}
\pgfplotsset{compat=1.18}

@@

These spaces are examples of Lie groups, formally defined as a smooth manifold equipped with a group structure.
They have applications in physics, robotics, stochastic processes, information geometry, and many other areas see~\cite{Chirikjian:2009,Chirikjian:2012},
but are also interesting from their mathematical viewpoint~\cite{HilgertNeeb:2012}
and their numerical aspects, for example when solving differential equations on Lie groups~\cite{Munthe-Kaas:1998,IserlesMunthe-KaasNoersettZanna:2000}.

@@

This paper provides an overview of the main features of \verb|LieGroups.jl| as of version \LieGroupsVersion{}, whose logo is shown on the right.
%
After introducing some mathematical background in Section~\ref{sec:Notation},
we present the interface in Section~\ref{sec:Interface} for the Lie group, the Lie algebra, and group actions.
Section~\ref{sec:LieGroups} provides an overview of all currently implemented Lie groups.
Finally, in Section~\ref{sec:Example}, we demonstrate how to get started and use \verb|LieGroups.jl|.

@@

\begin{figure}[tbp]
    \centering
    \includegraphics[width=0.12\textwidth]{logo.png}
    \vspace{.5\baselineskip}
    \caption*{Logo of \texttt{LieGroups.jl}.}%
    \label{fig:liegroups_logo}
\end{figure}

@@

\subsection{Tangent spaces and the Lie algebra} The tangent space at the identity element \(e \in \mathcal{G}\), denoted by \(\mathfrak{g} = T_e\mathcal{G}\), plays a special role and is called the \term{Lie algebra} of the Lie group \(\mathcal{G}\).
The reason is that to represent arbitrary tangent vectors $X \in T_g\mathcal{G}$ at a point $g \in \mathcal{G}$ we can use the group operation: we denote by $\lambda_g(h) = g \cdot h$ the left multiplication with $g, h \in \mathcal{G}$.
For a map \(f\colon \mathcal G \to \mathcal H\) between two Lie groups, its \term{differential} (or push forward) at a point \(g \in \mathcal G\) is a linear map \(Df(g)\colon T_g\mathcal G \to T_{f(g)}\mathcal H\) between the corresponding tangent spaces.
Using the differential $D\lambda_g(h)\colon T_h\mathcal G \to T_{g\cdot h}\mathcal G$, we can generate a so-called \term{left-invariant vector field} $\mathcal X(g) \coloneqq D\lambda_g(e)[X]$ which is uniquely determined by the choice of $X \in \mathfrak{g}$.%
\footnote{Analogously, one can use the right multiplication $\rho_g(h) = h \cdot g$ and its differential $D\rho_g(h)$ to define right-invariant vector fields. \texttt{LieGroups.jl} uses left-invariant vector fields as default.}%
Hence we can identify tangent vectors $\mathcal X(g) \in T_g\mathcal G$ at arbitrary points $g \in \mathcal G$ with $X$ from the Lie algebra $\mathfrak{g}$.
\\
As a first example, consider $\mathcal G = (\mathbb R^n, +)$, where the tangent space at any point $g$ is again $\mathbb R^n$, especially at the identity $e=0$. We further have $\lambda_g(h) = g + h$ and hence $D\lambda_g(h)[X] = X$ for all $g, h, X \in \mathbb R^n$.
Here, a tangent vector $X$ induces the constraint vector field $\mathcal X(g) = X$ for all $g \in \mathbb R^n$.\\
As a second example, consider again the special orthogonal group $\mathrm{SO}(n)$.
The tangent space at the identity element $e = I$ is given by the Lie algebra $\mathfrak{so}(n) = T_e\mathrm{SO}(n) = \{ X \in \mathbb{R}^{n \times n} \mid X = -X^{\mathrm{T}} \}$ consists of all skew-symmetric $n \times n$ matrices.
For $g, h \in \mathrm{SO}(n)$ and $X \in \mathfrak{so}(n)$ we have $\lambda_g(h) = g \cdot h$ and hence $D\lambda_g(h)[X] = g \cdot X$. Here, the tangent vector $X \in \mathfrak{so}(n)$ induces the left-invariant vector field $\mathcal X(g) = g \cdot X \in T_g\mathrm{SO}(n)$, $g \in \mathrm{SO}(n)$.
In other words, this formulation allows to represent tangent vectors $Y\in T_g\mathrm{SO}(n)$ also using $X = g^{-1}Y \in \mathfrak{g}$.

An important tool to “move around” on the Lie group is the exponential map.
The (Lie group) exponential function \(\exp\colon \mathfrak{g} \to \mathcal{G}\) maps elements from the Lie algebra to the Lie group, and is formally defined \cite[Def.~9.2.2]{HilgertNeeb:2012} by evaluating the unique curve \(\gamma\colon \mathbb{R} \to \mathcal{G}\) that fulfils the differential equation
\begin{equation*}
\gamma'(t) = D\lambda_{\gamma(t)}(e)[X]
\quad\text{ with }\quad\gamma(0) = e \text{ and }\gamma'(0) = X
\end{equation*}
at time $t=1$, that is \(\exp(X) = \gamma(1)\).
This can be interpreted as starting at the identity element \(e \in \mathcal{G}\) and following the curve whose velocity at each point is given by the left-invariant vector field induced by \(X \in \mathfrak{g}\) for one time unit.

For \((\mathbb{R}^n, +)\) the exponential map is given by \(\exp(X) = X\), $X\in \mathbb R^n$ and for \(\mathrm{SO}(n)\) it is given by the matrix exponential \(\exp(X) = \e^{X}\) for all \(X \in \mathfrak{so}(n)\).
Additionally for the unit circle in the complex plane, using the group operation of multiplication, the exponential map is given by the complex exponential \(\exp(X) = \e^{\im X}\) for all \(X \in \mathbb{R}\).

@@

\begin{table}
    \centering
    \caption{Group operations available in \texttt{LieGroups.jl}.}
    \begin{tabular}{@{}ll@{}}
        \toprule
        \textbf{Group operation} $\cdot$ & comment/code\\
        \midrule
        \verb|AdditionGroupOperation| & falls back to \verb|+|\\
        \verb|AbelianMultiplicationGroupOperation| & falls back to \verb|*|\\
        \verb|MatrixMultiplicationGroupOperation| & falls back to \verb|*|\\
        \verb|PowerGroupOperation{Op}| & \verb|PowerLieGroup|\\
        \verb|ProductGroupOperation{Ops}| & product Lie groups\\
        \verb|SemidirectProductGroupOperation| & semidirect products\\
        \bottomrule
    \end{tabular}
    \label{tab:GroupOperations}
\end{table}

@@

In \verb|LieGroups.jl|, the group exponential function is given by \verb|exp(G,X)|%
%\footnote{in-place variant \texttt{exp!(G, g, X)}.}
.
If we want to “start walking” from another point, we can “move” (or interpret $X$) as being from the tangent space at some point $g$ and obtain by the chain rule the \term{exponential function} $\exp_g\colon\mathfrak g\to\mathcal G$ defined by $\exp_g(X)\coloneqq g \cdot \exp(X)$.
In \verb|LieGroups.jl|, the exponential map is implemented as \verb|exp(G, g, X)|%
%\footnote{in-place variant \texttt{exp!(G, h, g, X)}.}
.
As a word of carefulness, note that on the underlying manifold, there is a further exponential map, the Riemannian exponential map. The Riemannian exponential map \verb|exp(M, p, X)|
%\footnote{with the same in-place signature as before \texttt{exp!(M, q, p, X)}.}
can be distinguished in that the first argument is a manifold, and the tangent vector \verb|X| has to be from the tangent space at the point \verb|p| on the manifold.
Here, again, to access the Riemannian exponential, one can use the base manifold of the Lie group, i.e.\ \verb|exp(base_manifold(G), g, X)| and have to make sure that \verb|X| is from the tangent space at \verb|g|.

@@

Similar to points on the Lie group, when representing elements $X \in \mathfrak g$ from a Lie algebra, we do not type the general functions of the interface.
This allows to use either plain arrays or own structures to represent these in code.
\footnote{The optional abstract supertype \texttt{AbstractLieAlgebraTangentVector}, which is a \texttt{AbstractTangentVector}, is provided to also distinguish different representations here.}
Keep in mind, that the Lie algebra is a vector space, so that addition and scalar multiplication are assumed to be defined in case you use an individual data type.

@@

\subsubsection*{Topological basics.}
Given a Lie group \verb|G|, we obtain the Lie algebra by calling $\mathfrak g$ \verb|= LieAlgebra(G)|.
To access the Lie group again, use \verb|base_lie_group(|$\mathfrak g$\verb|)|. Similarly \verb|base_manifold(|$\mathfrak g$\verb|)| returns the underlying manifold of the Lie group.\\
As a technical detail, the Lie algebra is modelled as a tangent space\footnote{see the documentation of~\href{https://juliamanifolds.github.io/ManifoldsBase.jl/stable/metamanifolds/\#ManifoldsBase.TangentSpace}{ManifoldsBase.TangentSpace}} using the already mentioned \verb|Identity(G)| as base point.
A zero vector is generated via \verb|zero_vector(|$\mathfrak g$\verb|)| for the default representation and \verb|zero_vector(|$\mathfrak g$\verb|, T)| for a specific representation type \verb|T|%
%\footnote{the in-place variant \texttt{zero\_vector!($\mathfrak g$, X::T)} can infer the type \texttt{T} of \texttt{X} automatically.}
.

@@

\begin{table}
    \centering
    \caption{Group actions types available in \texttt{LieGroups.jl}.}
    \begin{tabular}{@{}ll@{}}
        \toprule
        \textbf{Group action type} & comment/code\\
        \midrule
        \verb|AdditionGroupAction| & falls back to \verb|+|\\
        \verb|ColumnwiseGroupAction{A}| & of a group action type \verb|A|\\
        \verb|(Inverse)LeftGroupOperationAction| & requires $\mathcal M = \mathcal G$\\
        \verb|LeftMultiplicationAction| & falls back to \verb|*|\\
        \verb|(Inverse)RightGroupOperationAction| & requires $\mathcal M = \mathcal G$\\
        \verb|RotationAroundAxisAction| & \\
        \verb|RowwiseGroupAction{A}| & of a group action type \verb|A|\\
        \bottomrule
    \end{tabular}
    \label{tab:GroupActionTypes}
\end{table}

@@

\begin{table}[tbp]
    \centering
    \caption{Implemented Lie groups in \texttt{LieGroups.jl} version~\LieGroupsVersion{}.}
    \begin{tabular}{@{}lll@{}}
        \toprule
        \textbf{Lie Group $\mathcal G$} & \textbf{Symbol} & comment/code\\
        \midrule
        \verb|CircleGroup()| & \(\mathbb{S}^1\) & 3 representations \\
        \verb|GeneralLinearGroup(n, F)| & \(\mathrm{GL}(n, \mathbb{F})\) & $\mathbb{F} \in \{\mathbb{R}, \mathbb{C}\}$\\
        \verb|HeisenbergGroup(n)| & \(\mathrm{H}(n)\)\\
        \verb|OrthogonalGroup(n)| & \(\mathrm{O}(n)\) &\\
        \midrule
        \verb|PowerLieGroup(G, n)| & \(\mathcal G^n\) & \verb|G^n|\\
        \verb|ProductLieGroup(G1, G2,...)| & \(\mathcal G_1 \times \mathcal G_2 \times \ldots\) & \verb|G1| $\times$ \verb|G2| $\times$ \ldots\\
        Semidirect product group & \(\mathcal G_1 \ltimes \mathcal G_2\) & \verb|G1| $\ltimes$ \verb|G2|\\
                                 & \(\mathcal G_1 \rtimes \mathcal G_2\) & \verb|G1| $\rtimes$ \verb|G2|\\
        \midrule
        \verb|SpecialEuclideanGroup(n)| & \(\mathrm{SE}(n)\) & \\
        \verb|SpecialGalileanGroup(n)| & \(\mathrm{SGal}(n)\) &  \\
        \verb|SpecialLinearGroup(n, F)| & \(\mathrm{SL}(n, \mathbb{F})\) & $\mathbb{F} \in \{\mathbb{R}, \mathbb{C}\}$\\
        \verb|SpecialOrthogonalGroup(n)| & \(\mathrm{SO}(n)\) &  \\
        \verb|SpecialUnitaryGroup(n)| & \(\mathrm{SU}(n)\) & \\
        \midrule
        \verb|SymplecticGroup(n)| & \(\mathrm{Sp}(2n)\) & \\
        \verb|TranslationGroup(n; field=|$\mathbb{F}$\verb|)| & \(\mathbb{F}^n\) & $\mathbb{F} \in \{\mathbb{R}, \mathbb{C}, \mathbb{H}\}$\\
        \verb|UnitaryGroup(n)| & \(\mathrm{U}(n)\) & \\
        \verb|ValidationLieGroup(G)| & \multicolumn{2}{l}{wraps \texttt{G} for numerical verification}\\
        \bottomrule
    \end{tabular}
    \label{tab:available_lie_groups}
\end{table}

@@

Given two Lie groups \(\mathcal{G} = (\mathcal M, \ast) \) and \(\mathcal{H} = (\mathcal N, \diamond)\), their (direct) product \(\mathcal{G} \times \mathcal{H}\) is again a Lie group \((\mathcal{M} \times \mathcal{N}, \cdot)\) with group operation defined component-wise, i.e.
\begin{equation*}
    (g_1, h_1) \cdot (g_2, h_2) = (g_1 \ast g_2, h_1 \diamond h_2)
\end{equation*}
for \(g_1, g_2 \in \mathcal{G}\) and \(h_1, h_2 \in \mathcal{H}\).
Since $\times$ is a binary operator in Julia, given two Lie Groups \verb|G,H|, their product Lie group can be constructed via \verb|G|\,$\times$\,\verb|H|.
The same applies for more than two groups, i.e. you can construct arbitrary long product Lie groups \verb|G1|\,$\times$\,\verb|G2|\,$\times$\,\verb|G3|\,$\times$\,\ldots.

@@

Note that inheriting from the power manifold, there are two different data types to represent data, per default a single large array. Alternatively, one can use the constructor \verb|PowerLieGroup(G, NestedPowerRepresentation(), n)| to represent points and tangent vectors as a vector of points.

@@

For the \verb|CircleGroup()|, three different representations of points are available: as angles in \([0, 2\pi)\), as complex numbers with unit norm, and as points on the unit circle in \(\mathbb{R}^2\).
These are constructed by specifying the representation of points, i.e. \verb|CircleGroup(|$\mathbb R$\verb|)|, \verb|CircleGroup(|$\mathbb C$\verb|)|, and \verb|CircleGroup(|$\mathbb R^2$\verb|)|, respectively. Note that each of these representations uses a different underlying manifold, namely \verb|Circle(|$\mathbb R$\verb|)|, \verb|Circle(|$\mathbb C$\verb|)|, and \verb|Sphere(1)|, respectively.
This is a case, where it is not only the representation of points and tangent vectors that differ, but indeed even the underlying manifold.

@@

    \begin{tikzpicture}
        \begin{axis}[
            grid=major,
            width=.5\textwidth,
            legend style={at={(0.95,0.9)}},
            xmin=-0.05, xmax=1.05,
            ymin=-0.525, ymax=0.625,
            unit vector ratio=1 1,
            title style={yshift=-.5\baselineskip,},
            title={Two Bézier curves on different Lie groups},
        ]
            % Control points
            % Quiver plot from CSV file
            \addplot[
                quiver={u=\thisrow{u}, v=\thisrow{v}, scale arrows=\arrowscale},
                -stealth, very thick, TolMutedBlue,
            ] table[col sep=comma] {data/bezier_example_control_points.csv};
            \addlegendentry{Control points}
            % Plot the base points from CSV file
            \addplot[only marks, mark=*, mark size=2pt, TolMutedBlue, forget plot,
                ] table[col sep=comma] {data/bezier_example_control_points.csv};
            % SO2xR2 Curve and quiver
            \addplot[solid, thick, TolMutedSand, forget plot] table[col sep=comma, x=x, y=y] {data/bezier_example_so2r2_dense_curve.csv};
            % Quiver plot from CSV file
            \addplot[
                quiver={u=\thisrow{u}, v=\thisrow{v}, scale arrows=\arrowscale},
                -stealth, TolMutedSand,
            ] table[col sep=comma] {data/bezier_example_so2r2.csv};
            \addlegendentry{$\mathrm{SO}(2)\times \mathcal T(2)$}
            % Plot the base points from CSV file
            \addplot[only marks, mark=*, mark size=1.5pt, TolMutedSand, forget plot,
                ] table[col sep=comma] {data/bezier_example_so2r2.csv};
            % SE2 Curve and quiver
            \addplot[solid, thick, TolMutedTeal, forget plot] table[col sep=comma, x=x, y=y] {data/bezier_example_se2_dense_curve.csv};
            % Quiver plot from CSV file
            \addplot[
                quiver={u=\thisrow{u}, v=\thisrow{v}, scale arrows=\arrowscale},
                -stealth, TolMutedTeal,
            ] table[col sep=comma] {data/bezier_example_se2.csv};
            \addlegendentry{$\mathrm{SE}(2)$}
            % Plot the base points from CSV file
            \addplot[only marks, mark=*, mark size=1.5pt, TolMutedTeal, forget plot,
                ] table[col sep=comma] {data/bezier_example_se2.csv};
        \end{axis}
    \end{tikzpicture}
    \caption{Bézier curves on the Lie groups \(\mathrm{SE}(2)\) (teal) and \(\mathrm{SO}(2) \times \mathbb{R}^2\) (sand) based on the same control points
    $g_1,\ldots,g_4$ (blue).}
    \label{fig:Bezier}
\end{figure}

@@

    }

>>>>

\documentclass{juliacon}
% JuliaCon proceedings template
\setcounter{page}{1}
\usepackage{amsmath,booktabs,mathtools,amssymb}
\usepackage{tikz}
\usetikzlibrary{arrows.meta}
\usepackage{pgfplots}
\usepackage{pgfplotstable}
\usepackage{balance}
\pgfplotsset{compat=1.18}

@@

These are examples of Lie groups, formally defined as a smooth manifold equipped with a group structure.
They have applications in physics, robotics, stochastic processes, information geometry, and many other areas see~\cite{Chirikjian:2009,Chirikjian:2012},
but are also interesting from their mathematical viewpoint~\cite{HilgertNeeb:2012}
and their numerical aspects, for example when solving differential equations on Lie groups~\cite{IserlesMunthe-KaasNoersettZanna:2000, Munthe-Kaas:1998}.

@@

This paper provides an overview of the main features of \verb|LieGroups.jl|, whose logo is shown on the right, as of version \LieGroupsVersion{}.
%
After introducing some mathematical background in Section~\ref{sec:Notation},
we present the interface in Section~\ref{sec:Interface} for the Lie group, the Lie algebra, and group actions.
Section~\ref{sec:LieGroups} provides an overview of all currently implemented Lie groups.
Finally, in Section~\ref{sec:Example}, we demonstrate how to get started and use \verb|LieGroups.jl|.

@@

\begin{figure}[tbp]
    \centering
    \includegraphics[width=0.15\textwidth]{logo.png}
    \vspace{.5\baselineskip}
    \caption*{Logo of \texttt{LieGroups.jl}.}%
    \label{fig:liegroups_logo}
\end{figure}

@@

\subsection{The Lie algebra} The tangent space at the identity element \(e \in \mathcal{G}\), denoted by \(\mathfrak{g} = T_e\mathcal{G}\), plays a special role and is called the \term{Lie algebra} $\frak g$ of the Lie group \(\mathcal{G}\).
The reason is that to represent arbitrary tangent vectors $X \in T_g\mathcal{G}$ at a point $g \in \mathcal{G}$ we can use the group operation: we denote by $\lambda_g(h) = g \cdot h$ the left multiplication with $g, h \in \mathcal{G}$.
For a map \(f\colon \mathcal G \to \mathcal H\) between two Lie groups, its \term{differential} (or push forward) at a point \(g \in \mathcal G\) is a linear map \(Df(g)\colon T_g\mathcal G \to T_{f(g)}\mathcal H\) between the corresponding tangent spaces.
Using the differential $D\lambda_g(h)\colon T_h\mathcal G \to T_{g\cdot h}\mathcal G$, we can generate a so-called \term{left-invariant vector field} defined by $\mathcal X(g) \coloneqq D\lambda_g(e)[X]$ which is uniquely determined by the choice of $X \in \mathfrak{g}$%
\footnote{Analogously, one can use the right multiplication $\rho_g(h) = h \cdot g$ and its differential $D\rho_g(e)$ to define right-invariant vector fields. \texttt{LieGroups.jl} uses left-invariant vector fields as default.}%
.
Hence we can identify tangent vectors $\mathcal X(g) \in T_g\mathcal G$ at arbitrary points $g \in \mathcal G$ with $X$ from the Lie algebra $\mathfrak{g}$.
\\
As a first example, consider $\mathcal G = (\mathbb R^n, +)$, where the tangent space at any point $g$ is again $\mathbb R^n$, especially at the identity $e=0$. We further have $\lambda_g(h) = g + h$ and hence $D\lambda_g(h)[X] = X$ for all $g, h, X \in \mathbb R^n$.
Here, a tangent vector $X$ induces the constraint vector field $\mathcal X(g) = X$ for all $g \in \mathbb R^n$.\\
As a second example, consider again the special orthogonal group $\mathrm{SO}(n)$.
The tangent space at the identity element $e = I$ is given by the Lie algebra $\mathfrak{so}(n) = T_e\mathrm{SO}(n) = \{ X \in \mathbb{R}^{n \times n} \mid X = -X^{\mathrm{T}} \}$ consists of all skew-symmetric $n \times n$ matrices.
For $g, h \in \mathrm{SO}(n)$ and $X \in \mathfrak{so}(n)$ we have $\lambda_g(h) = g \cdot h$ and hence $D\lambda_g(h)[X] = g \cdot X$. Here, the tangent vector $X \in \mathfrak{so}(n)$ induces the left-invariant vector field $\mathcal X(g) = g \cdot X \in T_g\mathrm{SO}(n)$, $g \in \mathrm{SO}(n)$.
In other words, this formulation allows to represent tangent vectors $Y\in T_g\mathrm{SO}(n)$ also using $X = g^{-1}Y \in \mathfrak{g}$.

An important tool to “move around” on the Lie group is the exponential.
The (Lie group) exponential (function) \(\exp\colon \mathfrak{g} \to \mathcal{G}\) maps elements from the Lie algebra to the Lie group, and is formally defined~\cite[Def.~9.2.2]{HilgertNeeb:2012} by evaluating the unique curve \(\gamma\colon \mathbb{R} \to \mathcal{G}\) that fulfils the differential equation
\begin{equation*}
\gamma'(t) = D\lambda_{\gamma(t)}(e)[X]
\quad\text{ with }\quad\gamma(0) = e \text{ and }\gamma'(0) = X
\end{equation*}
at time $t=1$, that is \(\exp(X) = \gamma(1)\).
This can be interpreted as starting at the identity element \(e \in \mathcal{G}\) and following the curve whose velocity at each point is given by the left-invariant vector field induced by \(X \in \mathfrak{g}\) for one time unit.

For \((\mathbb{R}^n, +)\) the exponential is given by \(\exp(X) = X\), $X\in \mathbb R^n$ and for \(\mathrm{SO}(n)\) it is given by the matrix exponential \(\exp(X) = \e^{X}\) for all \(X \in \mathfrak{so}(n)\).
Additionally for the unit circle in the complex plane, using the group operation of multiplication, the exponential function is given by the complex exponential \(\exp(X) = \e^{\im X}\) for all \(X \in \mathbb{R}\).

From the interpretation of the exponential function to follow a curve starting at the identity, we can define the (Lie group) \term{exponential map} to “start from an arbitrary point” \(g \in \mathcal{G}\) by \(\exp_g\colon \mathfrak{g} \to \mathcal{G}\), \(\exp_g(X) \coloneqq g \cdot \exp(X)\) for all \(X \in \mathfrak{g}\).

@@

\begin{table}
    \centering
    \caption{Group operations available in \texttt{LieGroups.jl}.}
    \begin{tabular}{@{}ll@{}}
        \toprule
        \textbf{Group operation} $\cdot$ & comment/code\\
        \midrule
        \verb|AdditionGroupOperation| & falls back to \verb|+|\\
        \verb|AbelianMultiplicationGroupOperation| & falls back to \verb|*|\\
        \verb|MatrixMultiplicationGroupOperation| & falls back to \verb|*|\\
        \verb|PowerGroupOperation{Op}| & \verb|PowerLieGroup|\\
        \verb|ProductGroupOperation{Ops}| & product Lie groups\\
        \verb|SemidirectProductGroupOperation| & semidirect products\\
        \bottomrule
    \end{tabular}%
    \label{tab:GroupOperations}
\end{table}

@@

In \verb|LieGroups.jl|, the group exponential function is given by \verb|exp(G, X)|%
%\footnote{in-place variant \texttt{exp!(G, g, X)}.}
.
If we want to “start walking” from another point, we can “move” (or interpret $X$) as being from the tangent space at some point $g$ and obtain by the chain rule the exponential map $\exp_g\colon\mathfrak g\to\mathcal G$ defined by $\exp_g(X)\coloneqq g \cdot \exp(X)$.
In \verb|LieGroups.jl|, the exponential map is implemented as \verb|exp(G, g, X)|%
%\footnote{in-place variant \texttt{exp!(G, h, g, X)}.}
.
Note that on the underlying manifold, there is a further exponential map, the Riemannian exponential map.
This usually differs from the Lie group exponential map.
The Riemannian exponential map \verb|exp(M, p, X)|
%\footnote{with the same in-place signature as before \texttt{exp!(M, q, p, X)}.}
can be distinguished in that the first argument is a manifold, and the tangent vector \verb|X| has to be from the tangent space at the point \verb|p| on the manifold.
Here, again, to access the Riemannian exponential, one can use the base manifold of the Lie group, i.e.\ \verb|exp(base_manifold(G), g, X)| and have to make sure that \verb|X| is from the tangent space at \verb|g|.

@@

Similar to points on the Lie group, when representing elements $X \in \mathfrak g$ from a Lie algebra, we do not type the general functions of the interface.
This allows to use either plain arrays or own structures to represent these in code%
\footnote{The optional abstract supertype \texttt{AbstractLieAlgebraTangentVector}, which is a \texttt{AbstractTangentVector}, is provided to also distinguish different representations here.}%
.
Keep in mind, that the Lie algebra is a vector space, so that addition and scalar multiplication are assumed to be defined in case you use an individual data type.

@@

\subsubsection*{Topological basics}
Given a Lie group \verb|G|, we obtain the Lie algebra by calling $\mathfrak g$ \verb|= LieAlgebra(G)|.
To access the Lie group again, use \verb|base_lie_group(|$\mathfrak g$\verb|)|. Similarly \verb|base_manifold(|$\mathfrak g$\verb|)| returns the underlying manifold of the Lie group.\\
As a technical detail, the Lie algebra is modelled as a tangent space\footnote{see the documentation of~\href{https://juliamanifolds.github.io/ManifoldsBase.jl/stable/metamanifolds/\#ManifoldsBase.TangentSpace}{ManifoldsBase.TangentSpace}} using the already mentioned \verb|Identity(G)| as base point.
A zero vector is generated via \verb|zero_vector(|$\mathfrak g$\verb|)| for the default representation and \verb|zero_vector(|$\mathfrak g$\verb|, T)| for a specific representation type \verb|T|%
%\footnote{the in-place variant \texttt{zero\_vector!($\mathfrak g$, X::T)} can infer the type \texttt{T} of \texttt{X} automatically.}
.

@@

\begin{table}
    \centering
    \caption{Group actions types available in \texttt{LieGroups.jl}.}
    \begin{tabular}{@{}ll@{}}
        \toprule
        \textbf{Group action type} & comment/code\\
        \midrule
        \verb|AdditionGroupAction| & falls back to \verb|+|\\
        \verb|ColumnwiseGroupAction{A}| & of a group action type \verb|A|\\
        \verb|(Inverse)LeftGroupOperationAction| & requires $\mathcal M = \mathcal G$\\
        \verb|LeftMultiplicationAction| & falls back to \verb|*|\\
        \verb|(Inverse)RightGroupOperationAction| & requires $\mathcal M = \mathcal G$\\
        \verb|RotationAroundAxisAction| & \\
        \verb|RowwiseGroupAction{A}| & of a group action type \verb|A|\\
        \bottomrule
    \end{tabular}%
    \label{tab:GroupActionTypes}
\end{table}

@@

\begin{table}[tbp]
    \centering
    \caption{Implemented Lie groups in \texttt{LieGroups.jl} version~\LieGroupsVersion{}.}
    \begin{tabular}{@{}lll@{}}
        \toprule
        \textbf{Lie Group $\mathcal G$} & \textbf{Symbol} & comment/code\\
        \midrule
        \verb|CircleGroup()| & \(\mathbb{S}^1\) & 3 representations \\
        \verb|GeneralLinearGroup(n, F)| & \(\mathrm{GL}(n, \mathbb{F})\) & $\mathbb{F} \in \{\mathbb{R}, \mathbb{C}\}$\\
        \verb|HeisenbergGroup(n)| & \(\mathrm{H}(n)\)\\
        \verb|OrthogonalGroup(n)| & \(\mathrm{O}(n)\) &\\
        \midrule
        \verb|PowerLieGroup(G, n)| & \(\mathcal G^n\) & \verb|G^n|\\
        \verb|ProductLieGroup(G1, G2,...)| & \(\mathcal G_1 \times \mathcal G_2 \times \ldots\) & \verb|G1| $\times$ \verb|G2| $\times$ \ldots\\
        Semidirect product group & \(\mathcal G_1 \ltimes \mathcal G_2\) & \verb|G1| $\ltimes$ \verb|G2|\\
                                 & \(\mathcal G_1 \rtimes \mathcal G_2\) & \verb|G1| $\rtimes$ \verb|G2|\\
        \midrule
        \verb|SpecialEuclideanGroup(n)| & \(\mathrm{SE}(n)\) & \\
        \verb|SpecialGalileanGroup(n)| & \(\mathrm{SGal}(n)\) &  \\
        \verb|SpecialLinearGroup(n, F)| & \(\mathrm{SL}(n, \mathbb{F})\) & $\mathbb{F} \in \{\mathbb{R}, \mathbb{C}\}$\\
        \verb|SpecialOrthogonalGroup(n)| & \(\mathrm{SO}(n)\) &  \\
        \verb|SpecialUnitaryGroup(n)| & \(\mathrm{SU}(n)\) & \\
        \midrule
        \verb|SymplecticGroup(n)| & \(\mathrm{Sp}(2n)\) & \\
        \verb|TranslationGroup(n; field=|$\mathbb{F}$\verb|)| & \(\mathbb{F}^n\) & $\mathbb{F} \in \{\mathbb{R}, \mathbb{C}, \mathbb{H}\}$\\
        \verb|UnitaryGroup(n)| & \(\mathrm{U}(n)\) & \\
        \verb|ValidationLieGroup(G)| & \multicolumn{2}{l}{wraps \texttt{G} for numerical verification}\\
        \bottomrule
    \end{tabular}%
    \label{tab:available_lie_groups}
\end{table}

@@

Given two Lie groups \(\mathcal{G} = (\mathcal M, \ast) \) and \(\mathcal{H} = (\mathcal N, \diamond)\), their (direct) product \(\mathcal{G} \times \mathcal{H}\) is again a Lie group \((\mathcal{M} \times \mathcal{N}, \cdot)\) with group operation defined component-wise, i.e.
\begin{equation*}
    (g_1, h_1) \cdot (g_2, h_2) = (g_1 \ast g_2, h_1 \diamond h_2)
\end{equation*}
for \(g_1, g_2 \in \mathcal{G}\) and \(h_1, h_2 \in \mathcal{H}\).
Since $\times$ is a binary operator in Julia, given two Lie Groups \verb|G,H|, their product Lie group can be constructed via \verb|G|\,$\times$\,\verb|H|.
The same applies for more than two groups, i.e.\ you can construct arbitrary long product Lie groups \verb|G1|\,$\times$\,\verb|G2|\,$\times$\,\verb|G3|\,$\times$\,\ldots.

@@

Note that inheriting from the power manifold, there are two different data types to represent data, per default a single large array. Alternatively, one can use the constructor \verb|PowerLieGroup(G, NestedPowerRepresentation(), n)| to use the other representation of points and tangent vectors, namely as a vector of points/tangent vectors.

@@

For the \verb|CircleGroup()|, three different representations of points are available: as angles in \([0, 2\pi)\), as complex numbers with unit norm, and as points on the unit circle in \(\mathbb{R}^2\).
These are constructed by specifying the representation of points, i.e.\ \verb|CircleGroup(|$\mathbb R$\verb|)|, \verb|CircleGroup(|$\mathbb C$\verb|)|, and \verb|CircleGroup(|$\mathbb R^2$\verb|)|, respectively. Note that each of these representations uses a different underlying manifold, namely \verb|Circle(|$\mathbb R$\verb|)|, \verb|Circle(|$\mathbb C$\verb|)|, and \verb|Sphere(1)|, respectively.
This is a case, where it is not only the representation of points and tangent vectors that differ, but indeed even the underlying manifold.

@@

    \begin{tikzpicture}
        \begin{axis}[
            grid=major,
            width=.5\textwidth,
            legend style={at={(0.95,0.9)}},
            xmin=-0.05, xmax=1.05,
            ymin=-0.525, ymax=0.625,
            unit vector ratio=1 1,
            title style={yshift=-.5\baselineskip,},
            title={Two Bézier curves on different Lie groups},
        ]
            % Control points
            % Quiver plot from CSV file
            \addplot[
                quiver={u=\thisrow{u}, v=\thisrow{v}, scale arrows=\arrowscale},
                -stealth, very thick, TolMutedBlue,
            ] table[col sep=comma] {data/bezier_example_control_points.csv};
            \addlegendentry{Control points}
            % Plot the base points from CSV file
            \addplot[only marks, mark=*, mark size=2pt, TolMutedBlue, forget plot,
                ] table[col sep=comma] {data/bezier_example_control_points.csv};
            % SO2xR2 Curve and quiver
            \addplot[solid, thick, TolMutedSand, forget plot] table[col sep=comma, x=x, y=y] {data/bezier_example_so2r2_dense_curve.csv};
            % Quiver plot from CSV file
            \addplot[
                quiver={u=\thisrow{u}, v=\thisrow{v}, scale arrows=\arrowscale},
                -stealth, TolMutedSand,
            ] table[col sep=comma] {data/bezier_example_so2r2.csv};
            \addlegendentry{$\mathrm{SO}(2)\times \mathcal T(2)$}
            % Plot the base points from CSV file
            \addplot[only marks, mark=*, mark size=1.5pt, TolMutedSand, forget plot,
                ] table[col sep=comma] {data/bezier_example_so2r2.csv};
            % SE2 Curve and quiver
            \addplot[solid, thick, TolMutedTeal, forget plot] table[col sep=comma, x=x, y=y] {data/bezier_example_se2_dense_curve.csv};
            % Quiver plot from CSV file
            \addplot[
                quiver={u=\thisrow{u}, v=\thisrow{v}, scale arrows=\arrowscale},
                -stealth, TolMutedTeal,
            ] table[col sep=comma] {data/bezier_example_se2.csv};
            \addlegendentry{$\mathrm{SE}(2)$}
            % Plot the base points from CSV file
            \addplot[only marks, mark=*, mark size=1.5pt, TolMutedTeal, forget plot,
                ] table[col sep=comma] {data/bezier_example_se2.csv};
        \end{axis}
    \end{tikzpicture}
    \caption{Bézier curves on the Lie groups \(\mathrm{SE}(2)\) (teal) and \(\mathrm{SO}(2) \times \mathbb{R}^2\) (sand) based on the same control points
    $g_1,\ldots,g_4$ (blue).}%
    \label{fig:Bezier}
\end{figure}

@@

    }%
